Supporting Information: The Regional Effects of Marine Protected Areas

### 2020-04-22 documents/ovando-regional-effects-of-mpas-si-v2.Rmd
confirm knit of corrected supplemental information and paper

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)

@@

@@

@@

 sim_mpa_effects <- outcomes %>% 
  filter(set == "ci",
         final_depletion > 0,
         f_v_m > .5) %>% 
  mutate(year = years_protected + 2002) %>% 
  filter(year <= max(pisco_data$year)) %>% 
  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
  filter(year >= 2003,
         !is.na(binned_year)) %>% 
  mutate(mpa_effect = pmin(mpa_effect,3))
```

@@

```{r classcode}

@@

```{r did-coefs, results = "asis"}

@@

```{r mean_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}

@@

```{r no-bin-plot, fig.cap="For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins. For this run we estimate the effect annually, relative to the year 2000. Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = FALSE}
no_bin <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      iter = 2500,
      first_year = 2001,
      bin_width = 1,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data
    )
  )

@@

```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run. "}
mean_fit <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      fit_mean = TRUE,
      iter = 2500,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data
    )
  )

@@

 exclude zeros

@@

@@

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested") & !str_detect(variable, "site_side")) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x =variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group, labeller=labeller(group=c(
     seeing_non_nested_betas="Observation Model\n(seeing)",
     seen_non_nested_betas = "Observed Model\n(seen)"
   ))) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))

@@

 non_nested_site_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested") & str_detect(variable, "site_side")) %>%
  mutate(variable = tolower(variable)) %>% 
  mutate(variable = str_replace(variable, "site_side","")) %>% 
  mutate(island = str_replace(str_split(variable, '_', simplify = TRUE)[,1],'-','')
 ) %>% 
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_grid(island~group, scales = "free",labeller=labeller(group=c(
     seeing_non_nested_betas="Observation Model\n(seeing)",
     seen_non_nested_betas = "Observed Model\n(seen)"),
     island=c(anacapa="ANA",sci="SCI",smi="SMI",sri="SRI"))) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10)) + 
  theme_minimal()

@@

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group,labeller=labeller(group=c(
     seeing_region_cluster_betas="Observation Model\n(seeing)",
     seen_region_cluster_betas = "Observed Model\n(seen)")))

@@

targeted_trends <- report$zissou_estimates %>% 
  filter(str_detect(variable, "targeted_did")) %>% 
  group_by(variable) %>% 
  mutate(year =zissou_fit$did_data$year %>% unique(),
         labvar=ifelse(variable=="nontargeted_did_betas","Nontargeted","Targeted")) %>% 
  ungroup() %>% 
  ggplot() + 
  geom_line(aes(year, estimate, color = labvar),alpha = 0.5) +
  geom_pointrange(aes(year, estimate, ymin = lower, ymax = upper, color = labvar)) + 
  scale_color_d3(name = "") +
  theme(legend.position = c(0.8,0.9))

@@

fit_report <- base_run$tmb_fit[[1]]$zissou_report

observed_seen <- base_run$tmb_fit[[1]]$zissou_data$log_density

observed_seeing <- base_run$tmb_fit[[1]]$zissou_data$any_seen

@@

a = base_run$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
  as_data_frame() %>% 
  select(contains("mean_")) %>% 
  mutate(resid = seen_diagnostics$residuals) %>% 
  gather(variable, value, -resid) %>% 
  ggplot(aes(value, resid)) + 
  geom_point() + 
  facet_wrap(~variable)

@@

roc_plot <- rocplot + style_roc(theme = zissou_theme)

@@

@@

years <- unique(fitted_data$year)

years <- zissou_fit$did_data$year %>% unique()

abundance_trends <- data_frame(abundance_hat = fit_report$abundance_hat,
                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
  mutate(log_abundance_hat = log(abundance_hat)) %>%
  group_by(classcode) %>%
  mutate(year = years) %>%
  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
  ungroup() %>%
  rename(numeric_classcode = classcode) %>%
  left_join(life_history, by = "numeric_classcode" )

@@

numeric_cor_plot <- base_run$data[[1]] %>%
  select_if(is.numeric) %>%
  select(-log_density, -targeted) %>%
  na.omit() %>%
  corrr::correlate() %>%
  gather("colname", "correlation", -rowname) %>%
  ggplot(aes(rowname, colname, fill = correlation)) +
  geom_tile() +
  geom_text(aes(rowname, colname, label = round(correlation, 2))) +
  scale_fill_gradient2(
    low = "tomato",
    mid = "white",
    high = "steelblue",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1,
    size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

@@

#### Sensitivity to "missing" observations

Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study. If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly? The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation? For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect. If that species was never observed at that site, we do not include a zero for that species. Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled. However, this is clearly a strong assumption. For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data. To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e. not including any zeros in the data, (Fig.S\@ref(fig:nozero-raw-trend)). The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.

```{r nozero-raw-trend, fig.cap = "Centered and scaled mean annual density, excluding zeros, of included fishes (points) and smoothed means of targeted and non-targeted groups (line) over time" }

nozero_trend_plot <- base_run$data[[1]] %>% 
  filter(any_seen == T) %>% 
  select(year, targeted,classcode, log_density, any_seen) %>% 
  mutate(density = exp(log_density) * any_seen) %>% 
  group_by(classcode,targeted, year) %>% 
  summarise(md = mean(density)) %>% 
  group_by(classcode,targeted) %>% 
  mutate(smd = (md - mean(md)) / sd(md)) %>% 
  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.25) +
  geom_smooth() +
  scale_color_d3(name = "Targeted?") + 
  scale_fill_d3(name = "Targeted?") +
  labs(y = "Centered and Scaled Mean Density", x = "Year")

nozero_trend_plot
```

@@

```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}

trophic_groups <- cip_data %>% 
  select(classcode, broadtrophic) %>% 
  unique()

@@

| Variable | Distribution|
|---------:|:-----------:|
| Scientific Name | Drawn from all possible species in `FishLife` (@thorson2017c)|
|steepness (h) | ~uniform(0.6,0.95)|
|Adult movement ($\sigma_{s=a}$) | ~uniform(0,0.5 * P)|
|Larval movement ($\sigma_{s=l}$) | ~uniform(0,0.5 * P)|
|Recruitment variation ($\sigma_{r}$) |  $\in\{0,0.05,.1\}$|
|Recruitment autocorrelation ($ac_{r}$) |  $\in\{0,0.05,.1\}$|
|DD adult movement (dmod) | $\in\{0.25,1\}$|
|Density-dependence timing | $\in\{local, global,post-dispersal\}$|
|\% Patches in MPA | ~uniform(0.01,1)|
|Initial fishing relative to natural mortality | ~uniform(0.01,4)|
|Selectivity as a multiple of maturity length| ~uniform(0.1,1.25)|
|Fleet model| $\in\{open-access, constant-effort,constant-catch\}$|
|Spatial effort model| $\in\{uniform, biomass,profits\}$|
|Years into simulation to start MPA| ~round(uniform(5,0.66 * T))|
|MPA is sprinkler?| $\in\{TRUE,FALSE\}$|
|Randomly place MPA?| $\in\{TRUE,FALSE\}$|
|Fleet reaction to MPA| $\in\{concentrate, abandon-ship\}$|
|Patchiness | ~uniform(0.01,0.75)|
|MPA habitat factor| $\in\{1, 4\}$|

>>>>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 7, fig.asp = .75)

@@

tmb_runs <- read_rds( file.path(run_dir,"tmb_model_fits.rds"))

@@

filter_summary <- used_data %>% 
  select(classcode, targeted) %>% 
  unique() %>% 
  group_by(targeted) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
  spread(targeted, n)

@@

 sim_mpa_effects <- outcomes %>% 
  filter(set == "ci",
         final_depletion > 0,
         f_v_m > .5) %>% 
  mutate(year = years_protected + 2002) %>% 
  filter(year <= max(pisco_data$year)) %>% 
  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
  filter(year >= 2003,
         !is.na(binned_year)) %>% 
  mutate(mpa_effect = pmin(mpa_effect,3))

 filter_summary <- used_data %>% 
  select(classcode, targeted) %>% 
  unique() %>% 
  group_by(targeted) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
  spread(targeted, n)

@@

```{r classcode, eval = FALSE}

@@

```{r did-coefs, results = "asis",eval = FALSE}

@@

```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}

@@

```{r no-bin-plot, fig.cap="For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins. For this run we estimate the effect annually, relative to the year 2000. Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = TRUE}
no_bin <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      iter = 2500,
      first_year = 2001,
      bin_width = 1,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data
    )
  )

@@

```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run. "}
mean_fit <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      fit_mean = TRUE,
      iter = 2500,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data,
      refresh = 1000
    )
  )

@@

Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study. If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly? The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation? For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect. If that species was never observed at that site, we do not include a zero for that species. Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled. However, this is clearly a strong assumption. For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data. To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e. not including any zeros in the data, (Fig.S\@ref(fig:no-zero-fit)). The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.

@@

tmb_runs <- tmb_runs %>%
  mutate(tmb_fit = map(tmb_fit,"result")) %>%
  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
  mutate(did_plot = map(processed_fits, "did_plot"))

betas <- tmb_runs$processed_fits[[1]]$betas

@@

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested") & !str_detect(variable, "site_side")) %>%
  mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x =variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group, labeller=labeller(group=c(
     seeing_non_nested_betas="Observation Model\n(seeing)",
     seen_non_nested_betas = "Observed Model\n(seen)"
   ))) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))

@@

 non_nested_site_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested") & str_detect(variable, "site_side")) %>%
    mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
  mutate(variable = tolower(variable)) %>% 
  mutate(variable = str_replace(variable, "site_side","")) %>% 
  mutate(island = str_replace(str_split(variable, '_', simplify = TRUE)[,1],'-','')
 ) %>% 
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_grid(island~group, scales = "free",labeller=labeller(group=c(
     seeing_non_nested_betas="Observation Model\n(seeing)",
     seen_non_nested_betas = "Observed Model\n(seen)"),
     island=c(anacapa="ANA",sci="SCI",smi="SMI",sri="SRI"))) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10)) + 
  theme_minimal()

@@

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
    mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group,labeller=labeller(group=c(
     seeing_region_cluster_betas="Observation Model\n(seeing)",
     seen_region_cluster_betas = "Observed Model\n(seen)")))

@@

report <- tmb_runs$tmb_fit[[1]]

targeted_trends <- report$zissou_estimates %>% 
  filter(str_detect(variable, "targeted_did")) %>% 
  mutate(variable = str_remove_all(variable,"(?=\\.).*" )) %>% 
  group_by(variable) %>% 
  mutate(year =report$did_data$year %>% unique(),
         labvar=ifelse(variable=="nontargeted_did_betas","Nontargeted","Targeted")) %>% 
  ungroup() %>% 
  ggplot() + 
  geom_line(aes(year, estimate, color = labvar),alpha = 0.5) +
  geom_pointrange(aes(year, estimate, ymin = lower, ymax = upper, color = labvar)) + 
  scale_color_d3(name = "") +
  theme(legend.position = c(0.8,0.9))

@@

fit_report <- tmb_runs$tmb_fit[[1]]$zissou_report

observed_seen <- tmb_runs$tmb_fit[[1]]$zissou_data$log_density

observed_seeing <- tmb_runs$tmb_fit[[1]]$zissou_data$any_seen

@@

a = tmb_runs$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
  as_data_frame() %>% 
  select(contains("mean_")) %>% 
  mutate(resid = seen_diagnostics$residuals) %>% 
  gather(variable, value, -resid) %>% 
  ggplot(aes(value, resid)) + 
  geom_point() + 
  facet_wrap(~variable)

@@

roc_plot <- rocplot + style_roc()

@@

fitted_data <- used_data

@@

years <- report$did_data$year %>% unique()

abundance_trends <-
  data_frame(
    abundance_hat = fit_report$abundance_hat,
    classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))
  ) %>%
  mutate(log_abundance_hat = log(abundance_hat)) %>%
  group_by(classcode) %>%
  mutate(year = years) %>%
  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
  ungroup() %>%
  rename(numeric_classcode = classcode) %>%
  left_join(life_history, by = "numeric_classcode")

@@

numeric_cor_plot <- used_data %>%
  select_if(is.numeric) %>%
  select(contains("mean_"), n_obs) %>%
  na.omit() %>%
  corrr::correlate() %>%
  gather("colname", "correlation", -rowname) %>%
  ggplot(aes(rowname, colname, fill = correlation)) +
  geom_tile() +
  geom_text(aes(rowname, colname, label = round(correlation, 2))) +
  scale_fill_gradient2(
    low = "tomato",
    mid = "white",
    high = "steelblue",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1,
    size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

@@

@@

```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}
cip_data <- used_data
trophic_groups <- used_data %>% 
  select(classcode, broadtrophic) %>% 
  unique()

@@

| Variable | Distribution|
|---------:|:-----------:|
| Scientific Name | Drawn from all possible species in `FishLife` (@thorson2017c)|
|steepness (h) | ~uniform(0.6,0.95)|
|Adult movement ($\sigma_{s=a}$) | ~uniform(0,0.25 * P)|
|Larval movement ($\sigma_{s=l}$) | ~uniform(0,0.25 * P)|
|Recruitment variation ($\sigma_{r}$) |  $\in\{0,0.05,.1,.2\}$|
|Recruitment autocorrelation ($ac_{r}$) |  $\in\{0,0.05,.1,.2\}$|
|DD adult movement (dmod) | $\in\{0.25,1\}$|
|Density-dependence timing | $\in\{local, global,post-dispersal\}$|
|\% Patches in MPA | ~uniform(0.01,1)|
|Initial fishing relative to natural mortality | ~uniform(0.01,4)|
|Selectivity as a multiple of maturity length| ~uniform(0.1,1.25)|
|Fleet model| $\in\{open-access, constant-effort,constant-catch\}$|
|Spatial effort model| $\in\{uniform, biomass,profits\}$|
|Years into simulation to start MPA| ~round(uniform(5,0.66 * T))|
|MPA is sprinkler?| $\in\{TRUE,FALSE\}$|
|Randomly place MPA?| $\in\{TRUE,FALSE\}$|
|Fleet reaction to MPA| $\in\{concentrate, abandon-ship\}$|
|Patchiness | ~uniform(0.01,0.75)|
|MPA habitat factor| $\in\{1, 4\}$|
